\begin{thebibliography}{x}

	\bibitem{EMC} EMC corporation e IDC (international data corporation),
	\textit{``The digital universe of opportunities"}, 2014. \url{https://www.emc.com/collateral/analyst-reports/idc-digital-universe-2014.pdf}

	\bibitem{yole} M. Grao Txapartegi y Dr. Eric Mounier. 
	\textit{``Data Center Technologies. New Technologies and Architectures for Efficient Data Centers"}, Yole Développement, 2015. \url{http://www.i-micronews.com/images/Flyers/Power/Yole_New_Technologies_and_Architectures_for_Efficient_Data_Centers_Flyer_web.pdf}

	\bibitem{consumo1} EYP Mission Critical Facilities Inc., New York, 2011.

	\bibitem{consumo2} Rychard L. Sawyer, 
	\textit{``Calculating power requirements for data centers"}, APC y Schneider Electric, 2011. \url{http://www.apc.com/salestools/VAVR-5TDTEF/VAVR-5TDTEF_R1_EN.pdf}

	\bibitem{PUE} V. Avelar, D. Azevedo y A. French.
		\textit{``Pue: A comprehensive examination of the metric"}, 2014. \url{ https://datacenters.lbl.gov/sites/all/files/WP49-PUE\%20A\%20Comprehensive\%20Examination\%20of\%20the\%20Metric_v6.pdf}

	\bibitem{ITEfficiency} M. Zapater Sancho, P. Arroba García, J. M. Moya Fernández y Z. Bankovic. 
		\textit{``A state-of-the-art on energy eficiency in today's datacentres: Researcher's contributions and practical approaches"}, Vol. XII, Oct 2011. 

	\bibitem{DVFS} P. Chauhan y M. Gupta.
		\textit{``Energy Aware Cloud Computing Using Dynamic Voltage Frecuency Scalling"}, International Journal of Computer Science And Technology (IJCST), Vol 5, Issue 4, pp 195-199, Oct-Dec 2014. \url{ http://www.ijcst.com/vol54/2/49-Pooja-Chauhan.pdf}

	\bibitem{DPM} Silicon Valey Leadership Group,
		\textit{"Dynamic Power Management: Adjusting Data Center Capacity in Real-Time Energy Efficient Data Center Demonstration Project"}, 2009. \url{ http://svlg.org/wp-content/uploads/2012/12/PowerAssure_cs.pdf}

	\bibitem{VM} R. Buyya, A. Beloglazov y J. Abawajy. 
		\textit{``Energy-Efficient Management of Data Center Resources for Cloud Computing: A Vision, Architectural Elements, and Open Challenges"}, PDPTA, 2010. \url{https://pdfs.semanticscholar.org/65f3/92ea232878e416b077605427ff62d7fe0677.pdf}

	\bibitem{FreeCooling} Hainan Zhang, Shuangquan Shao, Hongbo Xu, Huiming Zou y Changqing Tian. 
	\textit{``Free cooling in a data centers: A review"}, Renewable and Sustainable Energy Reviews 35, pp 171-182, 2014. \url{ http://ac.els-cdn.com/S1364032114002445/1-s2.0-S1364032114002445-main.pdf?_tid=5780469c-30ef-11e7-bdaa-00000aab0f26&acdnat=1493919198_8346bfe1c779548bcc2e857ac03500bc} 

	\bibitem{OverCooling} Dan Hoffman.
		  \textit{``10 Techniques for improving data center power efficiency"}, Net app vision paper, Nov 2007. \url{http://www.uk.insight.com/content/dam/insight/EMEA/uk/shop/netapp/netappimproving-data-center-power-efficiency.pdf}

	\bibitem{cfugas} Y. Xie y W. L. Hung. 
		 \textit{``Temperature-aware task allocation and scheduling for embedded multiprocessor systems-on-chip (mpsoc) design"}, Journal of VLSI signal processing systems for signal, image and video technology, vol. 45, no. 3, pp. 177-189, 2006.

	\bibitem{GreenLSI} GreenLSI. \url{http://www.greenlsi.die.upm.es}

	\bibitem{RespConj} M. Zapater, J. L. Ayala y J. M. Moya. 
	           \textit{``Greendisc: a hw/sw energy optimization framework in globally distributed computation"}.  Ubiquitous Computing and Ambient Intelligence, pp 1-8, 2012. \url{http://oa.upm.es/16826/1/INVE_MEM_2012_137570.pdf}

	\bibitem{Monitorizacion} J. Pagan Ortiz, M. Zapater Sancho, O. Cubo, P. Arroba García, V. Martín Ayuso y J. M. Moya. 
	     \textit{A cyber-physical approach to combined hw-sw monitoring for improving energy eciency in data centers},  2013. \url{http://oa.upm.es/29886/1/INVE_MEM_2013_167142.pdf}

	\bibitem{Modelos} \textsc{M. Zapater}, \textsc{O. Tuncer}, \textsc{J. Ayala}, \textsc{J. Moya}, \textsc{K. Vaidyanathan}, \textsc{K. Gross} and \textsc{A. Coskun}. 
	      \textit{``Leakage-aware cooling management for improving server energy efficiency"}, IEEE Transactions on Parallel and Distributed Systems, vol 26, Issue:10, pp 2764-2777, oct 2015. 

	\bibitem{Esquema} Patricia Arroba, Jose M. Moya, Jose L. Ayala y Rajkumar Buyya.
	       \textit{``Proactive Power and Thermal Aware Optimizations for Energy-Efficient Cloud Computing"},  Design automation and test in europe (DATE), 2016. 

	\bibitem{control1} Katsuhiko Ogata. \textit{``Ingeniería de control moderna"}, Ed Pearson Education, 5º edición, 2010.

	\bibitem{control2} Benjamin C. Kuo. \textit{``Sistemas de control automático"}, Ed Prentice Hall, 7º Edición, 1996.
	       
	\bibitem{control3} Karl J. Astrom y Tore Hagglund. \textit{``Control PID avanzado"}, Ed Pearson Education, 2009. 

	\bibitem{PID} Bogdan M. Wilamoswski \& J. David Irwin. \textit{``The Industrial Electronics Handbook: Control and Mechatronics"}, 2º edition, CRC Press, 2011.
	
	\bibitem{fabricante1} Fabricante de controladores Omrom. \url{https://www.ia.omron.com/support/guide/53/introduction.html}

	\bibitem{fabricante2} Fabricante de controladores Omega. \url{http://www.omega.com/prodinfo/temperaturecontrollers.html}

	\bibitem{fabricante3} Fabricante de controladores Coulton. \url{http://www.coulton.com/QandA_controller}

	\bibitem{fabricante4} Fabricante de controladores Imopc. \url{http://www.imopc.es/content.php?p=spotlight_TempControl&lang=ES}

	\bibitem{Graphite1} Documentaci'on \textit{Graphite}. \url{http://graphite.readthedocs.io/en/latest/index.html}
	
	\bibitem{PIMOTE} Sara Álvarez Vinagre y Alfredo Tendero Casanova.
	\textit{Memoria proyecto Pimote SDG2}, GreenLSI, Departamento de Ingeniería Electrónica (DIE), Universidad Politécnica de Madrid, 2013.

           \bibitem{SPI} Protocolo de comunicación SPI. \url{http://www.i-micro.com/pdf/articulos/spi.pdf}

	\bibitem{curl} Librería \textit{curl} para peticiones HTTP en C. \url{https://curl.haxx.se/libcurl/}

	\bibitem{jansson} Librería \textit{jansson} para manejar objetos JSON en C. \url{https://jansson.readthedocs.io/en/2.10/}

\end{thebibliography}
